% ============================================================
% Section 6: Implementation and Deployment
% ============================================================
\section{Implementation and Deployment}
\label{sec:implementation}

\subsection{Software Stack}

The entire system is implemented in Python 3.8+ using only the standard library plus three external packages: \texttt{requests} (HTTP API calls), \texttt{numpy} (vector operations), and \texttt{mysql-connector-python} (database). No ML frameworks, GPU libraries, or deep learning toolkits are required. The codebase is approximately 1,800 lines of Python across 9 source files.

\subsection{Hermes Cron Pipeline}

The system runs as a Hermes Agent cron pipeline with four asynchronous stages:

\begin{itemize}
    \item \textbf{Market News Poller} (every 15 minutes, no-agent mode): \texttt{market\_news\_poller.py} --- polls the API, deduplicates via SHA-256 hash, writes new rows to \texttt{news\_events}. Runs as a stateless watchdog script; its stdout is delivered directly without LLM processing.
    \item \textbf{News Sentiment Scoring} (every 30 minutes, no-agent mode): \texttt{run\_news\_scoring.py} --- runs FinBERT lexicon scorer, statistical TF-IDF cluster lookup, and DeepSeek LLM scoring (up to 500 newest unscored headlines), then computes the ensemble combination and writes results to \texttt{sentiment\_signals}.
    \item \textbf{News Sentiment Calibration} (every 6 hours, no-agent mode): \texttt{run\_news\_calibrate.py} --- computes realized 24h price impacts, signal correlations, optimal ensemble weights, cluster maintenance (vocabulary and centroid rebuild from the last 7 days), and regime detection. Results are written to \texttt{calibration\_history}.
    \item \textbf{News Sentiment Pipeline} (every 3 hours, LLM-driven): \texttt{news-sentiment-pipeline.py} --- an orchestration script that runs ingest, scoring, and calibration stages in sequence with LLM oversight for logging and alerting.
\end{itemize}

\subsection{Impact Computation via Price Time Series}

The critical innovation in the implementation is the 24h impact computation. Rather than relying on the API's snapshot \texttt{pc\_esf} field (which reflects the change-percent at poller time, not the forward reaction), CRANE builds its own price time series from the historical \texttt{news\_events} rows. For each headline $h$ published at time $T$:

\begin{enumerate}
    \item Find the nearest ES futures price level \emph{before} $T$ (or at $T$ if no prior) from the time series sorted by \texttt{published\_at}.
    \item Find the nearest ES futures price level \emph{after} $T + 24$ hours.
    \item Compute $\text{impact}_{24h} = (\text{price}_{T+24} - \text{price}_T) / \text{price}_T \times 100$.
\end{enumerate}

This approach eliminates lookahead bias and provides a clean, deterministic ground truth for calibration. The \texttt{batch\_compute\_impacts()} function in \texttt{impact\_helpers.py} performs bulk binary-searched lookups for efficiency, computing impacts for up to 200 rows in 100--300ms.

\subsection{Live Sentiment Gauge Widget}
\label{sec:widget}

CRANE powers an interactive HTML sentiment gauge widget deployed on Tradeflags.com. The widget (Figure~\ref{fig:gauge}) provides:

\begin{itemize}
    \item \textbf{Main gauge}: A 180px circular gauge showing aggregate sentiment score on a $-100$ to $+100$ scale, with color coding (green = bullish, red = bearish, gray = neutral) and a dynamic arc fill.
    \item \textbf{Signal breakdown}: Three cards showing the current score from each signal (FinBERT lexicon, Statistical clusters, LLM) with their current ensemble weight percentages and confidence levels.
    \item \textbf{Multi-asset sentiment impacts}: Five mini-cards showing ES, NQ, CL, BTC, and ETH with name, price level, impact percentage (24h-forward), and a horizontal bar fill for visual comparison.
    \item \textbf{Market regime}: A bar indicator with color-coded dot showing the detected regime (bullish/bearish/neutral/volatile) from the calibration job.
    \item \textbf{Calibration info}: Samples count and current ensemble weights breakdown.
    \item \textbf{Theme tags}: The top-8 TF-IDF keywords automatically extracted from the most recent 50 headlines within the last 24 hours.
    \item \textbf{Auto-refresh}: The widget fetches a static JSON file generated by \texttt{sentiment\_api.py} (a RESTful query over \texttt{sentiment\_signals} and \texttt{calibration\_history}) every 60 seconds via JavaScript polling.
\end{itemize}

\begin{figure}[ht]
\centering
\small
\begin{tabular}{|c|}
\hline
\textbf{Live Gauge Preview (Dark Theme)} \\
\hline
\fbox{\parbox{0.85\textwidth}{\small
\vspace{0.3cm}
\textbf{Market Sentiment \hfill 5-Asset}

Regime: Bullish \quad Score: +12.3

\vspace{0.2cm}
\begin{tabular}{ccc}
\textbf{LLM} & \textbf{FinBERT} & \textbf{Statistical} \\
+8.5 (w:35\%) & +5.2 (w:20\%) & +16.1 (w:45\%)
\end{tabular}

\vspace{0.2cm}
\begin{tabular}{ccccc}
ES: +0.06\% & NQ: +0.15\% & CL: -0.37\% & BTC: -0.23\% & ETH: -0.31\%
\end{tabular}

\vspace{0.2cm}
\textbf{Themes}: tariff, ai, earnings, inflation, nvidia, oil, fed, gdp
\vspace{0.3cm}
}} \\
\hline
\end{tabular}
\caption{Live sentiment gauge mockup (dark theme, tradeflags.com style). The production widget fetches JSON from \texttt{sentiment\_api.py} and auto-refreshes every 60 seconds.}
\label{fig:gauge}
\end{figure}

\subsection{Source Code and Reproducibility}

The full source code and all analysis scripts are available in the project repository. The repository includes:

\begin{itemize}
    \item \texttt{ingest.py} --- API polling, deduplication, MySQL write
    \item \texttt{score.py} --- Three-signal scoring pipeline orchestration
    \item \texttt{calibrate.py} --- Auto-calibration, weight optimization, regime detection
    \item \texttt{utils/db.py} --- MySQL connection, schema creation, config loading
    \item \texttt{utils/finbert\_scorer.py} --- Financial lexicon with prospect-theory weighting
    \item \texttt{utils/stat\_scorer.py} --- TF-IDF cluster learner with multi-asset return buffers
    \item \texttt{utils/ensemble.py} --- Adaptive weighting, conflict detection
    \item \texttt{utils/llm\_scorer.py} --- DeepSeek zero-shot batch scoring
    \item \texttt{utils/impact\_helpers.py} --- 24h impact computation via price time series
    \item \texttt{sentiment\_api.py} --- RESTful JSON endpoint for the gauge widget
    \item \texttt{sentiment-gauge.html} --- Interactive dark-theme HTML gauge
    \item \texttt{docs/news-sentiment-engine-architecture.md} --- Full architecture documentation
\end{itemize}

All code is released under the MIT license.
